---
title: "GOASamplingDesign"
author: "Chris Rooper"
date: "3/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}

library(raster)
library(rgdal)
library(maptools)
library(maps)
library(gstat)
library(rgeos)
library(proj4)
library(sp)
library(mgcv)
library(tidyr)
library(arm)
library(PresenceAbsence)
library(ggplot2)
library(viridis)
library(dismo)
source("C:/Users/rooperc/Documents/Chris Work Stuff/R Software Help and Functions/Miscellaneous Functions/MappingExtras.R")

knitr::opts_chunk$set(echo = TRUE)
```

```{r hideme}

bathy<-raster("D:/GOA Coral and Sponge Model/Variables/RasterGrids/bathy_all")
bathy<-raster::aggregate(bathy,fact=100,fun=mean,filename="C:/Users/rooperc/Desktop/Deep Sea Coral Research/GOA Model Validation - Sigler/SampleDesign/RasterLayers/bathy",overwrite=TRUE)

coral<-raster("D:/GOA Coral and Sponge Model - Revisions/PredictionRasters/coralsENScpue")
coral<-raster::aggregate(coral,fact=100,fun=mean,filename="C:/Users/rooperc/Desktop/Deep Sea Coral Research/GOA Model Validation - Sigler/SampleDesign/RasterLayers/coral")
coralCV<-raster("D:/GOA Coral and Sponge Model - Revisions/PredictionRasters/coralsCVcpue")
coralCV<-raster::aggregate(coralCV,fact=100,fun=mean,filename="C:/Users/rooperc/Desktop/Deep Sea Coral Research/GOA Model Validation - Sigler/SampleDesign/RasterLayers/coralcv")

sponge<-raster("D:/GOA Coral and Sponge Model - Revisions/PredictionRasters/SpongeENScpue")
sponge<-raster::aggregate(sponge,fact=100,fun=mean,filename="C:/Users/rooperc/Desktop/Deep Sea Coral Research/GOA Model Validation - Sigler/SampleDesign/RasterLayers/sponge")
spongeCV<-raster("D:/GOA Coral and Sponge Model - Revisions/PredictionRasters/SpongeCVcpue")
spongeCV<-raster::aggregate(spongeCV,fact=100,fun=mean,filename="C:/Users/rooperc/Desktop/Deep Sea Coral Research/GOA Model Validation - Sigler/SampleDesign/RasterLayers/spongecv")

```





## Sample design

The goal of this code is to set up a few sample designs for validating a deep-sea coral model for the Gulf of Alaska.

```{r bring the pieces}

bathy<-raster("./RasterLayers/bathy")
coral<-raster("./RasterLayers/coral")
sponge<-raster("./RasterLayers/sponge")
coralCV<-raster("./RasterLayers/coralcv")
spongeCV<-raster("./RasterLayers/spongeCV")

plot(stack(bathy,coral,coralCV,sponge,spongeCV))

akland<-readOGR("./RasterLayers","AKland")
akland
plot(akland)

```

## Simple Random Sampling


```{r pressure, echo=FALSE}

srs_sites<-sampleRandom(bathy,size=300,xy=TRUE,sp=TRUE)
head(srs_sites)
dim(srs_sites)
hist(srs_sites$layer,breaks=10)

plot(bathy)
plot(akland,col="black",add=TRUE)
points(srs_sites,col="purple",pch=20)

srs_sites_ll<-coordinates(spTransform(srs_sites,crs("+proj=longlat +datum=WGS84")))


cpue<-raster::extract(coral,srs_sites)
hist(cpue,breaks=100)


```

## Stratified random sampling 

```{r stratified random sampling}

strata<-c(0,100,200,300,500,1000)
strata_raster<-cut(bathy,breaks=strata)
plot(strata_raster)

strata_sites<-raster::sampleStratified(strata_raster,size=300/5,xy=TRUE,sp=TRUE)
head(strata_sites)
dim(strata_sites)
hist(strata_sites$layer,breaks=10)

plot(bathy)
plot(akland,col="black",add=TRUE)
points(strata_sites,col="purple",pch=20)

strata_sites_ll<-coordinates(spTransform(sites,crs("+proj=longlat +datum=WGS84")))


cpue<-raster::extract(coral,strata_sites)
hist(cpue,breaks=100)


```
## Stratified random sampling with proportional allocation

nh=n*(Nh/N)

```{r stratified random proportional sampling}

strata<-c(0,100,200,300,500,1000)
strata_raster<-cut(bathy,breaks=strata)
strata_Nh<-data.frame(zonal(bathy,strata_raster,fun="count"))
strata_Nh

nh<-round(300*(strata_Nh$count)/(sum(strata_Nh$count)),0)

nh

pstrata_sites<-NULL

for(i in 1:length(nh)){
tmp<-strata_raster
  tmp[tmp!=i]<-NA
tmp1<-sampleRandom(tmp,size=nh[i],xy=TRUE)[,1:2]
pstrata_sites<-rbind(pstrata_sites,tmp1)}
pstrata_sites<-SpatialPoints(coords=pstrata_sites,proj4string=CRS(proj4string(bathy)))

head(pstrata_sites)
dim(pstrata_sites)
depth<-raster::extract(bathy,pstrata_sites)
hist(depth,breaks=10)

plot(bathy)
plot(akland,col="black",add=TRUE)
points(pstrata_sites,col="purple",pch=20)

pstrata_sites_ll<-spTransform(pstrata_sites,crs("+proj=longlat +datum=WGS84"))


cpue<-raster::extract(coral,pstrata_sites)
hist(cpue,breaks=100)

```


## Stratified random sampling with Neyman allocation 

Nh=n*Sh*Nh/(sum(Sh*Nh)
Nh = numberof samples in stratum h
Sh = standard deviation of stratum h

```{r stratified random neyman sampling}

strata<-c(0,100,200,300,500,1000)
strata_raster<-cut(bathy,breaks=strata)
strata_Sh<-data.frame(zonal(coral,strata_raster,fun="sd"))
strata_Sh
strata_Nh<-data.frame(zonal(coral,strata_raster,fun="count"))
strata_Nh

nh<-round(300*(strata_Sh$sd*strata_Nh$count)/(sum(strata_Sh$sd*strata_Nh$count)),0)
nh


nstrata_sites<-NULL

for(i in 1:length(nh)){
tmp<-strata_raster
  tmp[tmp!=i]<-NA
tmp1<-sampleRandom(tmp,size=nh[i],xy=TRUE)[,1:2]
nstrata_sites<-rbind(nstrata_sites,tmp1)}
nstrata_sites<-SpatialPoints(coords=nstrata_sites,proj4string=CRS(proj4string(bathy)))

head(nstrata_sites)
dim(nstrata_sites)
depth<-raster::extract(bathy,nstrata_sites)
hist(depth,breaks=10)

plot(bathy)
plot(akland,col="black",add=TRUE)
points(nstrata_sites,col="purple",pch=20)

nstrata_sites_ll<-spTransform(nstrata_sites,crs("+proj=longlat +datum=WGS84"))


cpue<-raster::extract(coral,nstrata_sites)
hist(cpue,breaks=100)

```

## Weighted sampling

Probability of a station being sampled is determined by its relative weight. In this case we can weight by CV of the model prediction or the model predicted CPUE itself.

```{r weighted random sampling}

wsites<-randomPoints(coral, n=300, prob=TRUE)

head(wsites)
dim(wsites)
depth<-raster::extract(bathy,wsites)
hist(depth,breaks=10)

plot(bathy)
plot(akland,col="black",add=TRUE)
points(wsites,col="purple",pch=20)

wsites<-SpatialPoints(coords=wsites,proj4string=CRS(proj4string(bathy)))
wsites_ll<-spTransform(wsites,crs("+proj=longlat +datum=WGS84"))


wcpue<-raster::extract(coral,wsites)
hist(wcpue,breaks=100)



```




